Implicit yaddayadda

### 2011-10-27 papers/PLDI2012/implicitization.tex
Fusion de l'intro de la section 3. et de la section Algorithm

\newcommand{\MM}[1]{{\color{blue} TODO(MM): #1}}
\newcommand{\JH}[1]{{\color{teal} TODO(JH): #1}}
\newcommand{\DM}[1]{{\color{violet} TODO(DM): #1}}

@@

We distinguish three phases in the main loop of the analysis:
\begin{enumerate}
\item Whe start finding a relevant subset $P \cup P'$ of the graph.
  Either the previous iteration or the initialization lead us to a
  state where there are no more paths in the previous subset $P$,
  starting at $p_i$, that make the abstract values of the successors
  grow (otherwise, the SMT solver would not have answered
  ``\emph{unsat}''). Narrowing iterations preserve this property.
  However, there may exist such paths in the entire multigraph, that
  are not in $P$. This phase computes these paths and adds them to
  $P'$ (line~\ref{alg=computeNewPaths} in
  algorithm~\ref{algo:combined}). We also use SMT solving to discover
  these new paths, but we slightly change the SMT formula given to the
  SMT solver: we now simply check for the satisfiability of $f(p_i)$
  instead of $f(p_i) \wedge P$. Since we already know that $f(p_i)
  \wedge P$ is unsatisfiable, all the paths given by the SMT solver
  won't be in $P$. This phase is described in~\ref{subsec:addingpaths}
  and corresponds to lines in~\ref{alg=start-add-paths} to
  \ref{alg=end-add-paths} algorithm~\ref{algo:combined}.
\item Given a new subset $P$, we search for paths starting at point
  $p_i \in P_R$, such that these paths are in $P$, i.e are included in
  the working subgraph. Each time we find a path, we update the
  abstract value of the destination point of the path. This is the
  phase explained in~\ref{subsec:ascending}, and corresponds to
  lines~\ref{alg=start-ascending} to~\ref{alg=end-ascending} in
  algorithm~\ref{algo:combined}.
\item We perform narrowing iterations the usual way
  (line~\ref{alg=narrowing} in algorithm~\ref{algo:combined}) and
  reiterate from step 1 unless there are no more states to explore.
\end{enumerate}

@@

The current working subset of paths, noted $P$, is stored using a
compact representation, such as binary decision diagrams.
We also maintain two sets of control points:
\begin{itemize}
	\item $A'$ contains the points in $P_R$ that may be the starting points of new
		feasible paths.
	\item $A$ contains the points in $P_R$ on which we apply the ascending iterations.
	Each time the abstract value of a control point $p$ is updated, $p$ is
	inserted in both $A$ and $A'$.
\end{itemize}

We first check which paths become feasible, so that we can include them into the
next subset of paths. The set $P'$ of new feasible paths is added to the
previous set $P$. Each starting points of paths in $P'$ is added into $A$.
We apply the path-focusing \cite{Monniaux_Gonnord_SAS11} to
the elements of $A$ until $A$ is empty. 
When $A$ becomes empty, it means the
invariant computation of the subgraph is finished. As proposed by
\citet{DBLP:conf/sas/GopanR07}, we can do some narrowing iterations.
These narrowing iterations allow to recover precision lost by
widening, before computing and taking into account new feasible paths.
Thus, our technique combines both the advantages of \emph{Guided Static
Analysis} and \emph{Path-focusing}.

After narrowing, we restart the analysis, starting from the computation the new
subgraph.

@@

@@

\subsection{Analysis algorithm}
Our tool operates over LLVM bitcode \citep{LLVM_langref,Lattner:2004:LCF:977395.977673}. LLVM bitcode is in \emph{single static assignment} (SSA) form: a given scalar variable is given a value at a single syntactic point in the program. In concrete terms, an assignment \lstinline|x=2*x+1;| gets translated into a definition $x_2 = 2x_1+1$, with distinct variables $x_1$ and $x_2$ corresponding to the same original variable \lstinline|x| at different points in the program.
LLVM makes it easy to follow definition-use and use-definition chains: for a given variable (say, $x_2$) one can immediately obtain its definition (say, $2x_1+1$).
One may see conversion to SSA form as a static precomputation of some of the symbolic propagations proposed by \citet{DBLP:conf/vmcai/Mine06} to enhance the precision of analyses.

>>>>

\newcommand{\MM}[1]{{\color{blue} TODO(MM): #1}}
\newcommand{\JH}[1]{{\color{teal} TODO(JH): #1}}
\newcommand{\DM}[1]{{\color{violet} TODO(DM): #1}}
% Pour les commentaires
\usepackage[utf8]{inputenc}

@@

@@

\MM{Il faudra peut-être une section ``détails d'implem'' où mettre
  l'histoire des BDDs pour P. Je laisse ici faute de mieux (peut-être
  que la section~\ref{sec:analysis-algorithm} ferait l'affaire ?),
  mais ça distrait un peu le lecteur}

The current working subset of paths, noted $P$ and initially empty, is
stored using a compact representation, such as binary decision
diagrams. We also maintain two sets of control points:
\begin{itemize}
	\item $A'$ contains the points in $P_R$ that may be the starting points of new
		feasible paths.
	\item $A$ contains the points in $P_R$ on which we apply the ascending iterations.
	Each time the abstract value of a control point $p$ is updated, $p$ is
	inserted in both $A$ and $A'$.
\end{itemize}

@@

We distinguish three phases in the main loop of the analysis:
\begin{enumerate}
\item \label{step:addingpaths} We start finding a new relevant subset
  $P \cup P'$ of the graph.
  Either the previous iteration or the initialization lead us to a
  state where there are no more paths in the previous subset $P$,
  starting at $p_i$, that make the abstract values of the successors
  grow (otherwise, the SMT solver would not have answered
  ``\emph{unsat}''). Narrowing iterations preserve this property.
  However, there may exist such paths in the entire multigraph, that
  are not in $P$. This phase computes these paths and adds them to
  $P'$ (line~\ref{alg=computeNewPaths} in
  algorithm~\ref{algo:combined}). We also use SMT solving to discover
  these new paths, but we slightly change the SMT formula given to the
  SMT solver: we now simply check for the satisfiability of $f(p_i)$
  instead of $f(p_i) \wedge P$. Since we already know that $f(p_i)
  \wedge P$ is unsatisfiable, all the paths given by the SMT solver
  won't be in $P$. This phase is described in~\ref{subsec:addingpaths}
  and corresponds to lines in~\ref{alg=start-add-paths} to
  \ref{alg=end-add-paths} algorithm~\ref{algo:combined}.
\item \label{step:ascending} Given a new subset $P$, we search for paths starting at point
  $p_i \in P_R$, such that these paths are in $P$, i.e are included in
  the working subgraph. Each time we find a path, we update the
  abstract value of the destination point of the path. This is the
  phase explained in~\ref{subsec:ascending}, and corresponds to
  lines~\ref{alg=start-ascending} to~\ref{alg=end-ascending} in
  algorithm~\ref{algo:combined}.
\item  \label{step:narrowing} We perform narrowing iterations the usual way
  (line~\ref{alg=narrowing} in algorithm~\ref{algo:combined}) and
  reiterate from step 1 unless there are no more points to explore,
  i.e. $A' = \emptyset$.
\end{enumerate}

Step~\ref{step:ascending} corresponds to the application of
path-focusing \cite{Monniaux_Gonnord_SAS11} to
the elements of $A$ until $A$ is empty. 
When $A$ becomes empty, it means the
invariant computation of the subgraph is finished. As proposed by
\citet{DBLP:conf/sas/GopanR07}, we can do some narrowing iterations.
These narrowing iterations allow to recover precision lost by
widening, \emph{before} computing and taking into account new feasible paths.
Thus, our technique combines both the advantages of \emph{Guided Static
Analysis} and \emph{Path-focusing}.

@@

\subsection{Analysis algorithm}
\label{sec:analysis-algorithm}
Our tool operates over LLVM bitcode \citep{LLVM_langref,Lattner:2004:LCF:977395.977673}. LLVM bitcode is in \emph{single static assignment} (SSA) form: a given scalar variable is given a value at a single syntactic point in the program. In concrete terms, an assignment \lstinline|x=2*x+1;| gets translated into a definition $x_2 = 2x_1+1$, with distinct variables $x_1$ and $x_2$ corresponding to the same original variable \lstinline|x| at different points in the program.
LLVM makes it easy to follow definition-use and use-definition chains: for a given variable (say, $x_2$) one can immediately obtain its definition (say, $2x_1+1$).
One may see conversion to SSA form as a static precomputation of some of the symbolic propagations proposed by \citet{DBLP:conf/vmcai/Mine06} to enhance the precision of analyses.
